---
title: "Maize"
author: "Adama NDOUR"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(gtsummary)
library(inspectdf)
library(sf)
```

## Import data
```{r}
df<- readxl::read_excel("input/maize.xls",sheet = "MaizeDec7-2023_final_no_IFDC")
```

## Overview of the data
```{r}
str(df)
```
## Data manipulation
```{r}
# Rename some variables
df<- df %>% rename_with(
   ~ c("all_tr1","recNP_tr8","diff_all","diff_recNP"),.cols= c("All (Trt1)","RecNP (Trt8)", "Yld. diff-All (%)", "Yld. diff-RecNP (%)"))

# Convert to numeric
df <- df %>% mutate(across(c("GYld_Kgpha","all_tr1","recNP_tr8"),
                           as.numeric))
```

## Inspect data
* Categorical variables
```{r}
df %>% select_if(is.character) %>% skim()
```
```{r}
df %>% inspect_cat()
```


```{r}
# numeric variables
df %>% select_if(is.numeric) %>% skim() %>% 
  knitr::kable()
```


## Descriptive statistics
* Yield per treatment
```{r message=FALSE, warning=FALSE}
df %>% select(c("Treat","GYld_Kgpha")) %>% 
  tbl_summary(
    by = Treat,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing_text = "(Missing)"
  ) %>% modify_spanning_header(all_stat_cols() ~ "**Treatments**") 
```

## Explory data analysis
* Map
```{r}
maize_areas<- st_as_sf(df %>% select(c("Lat_DD","Long_DD",
                                      "Treat","GYld_Kgpha","Region")) %>% drop_na() %>% filter(GYld_Kgpha>0,Treat==9),
                        coords=c("Long_DD","Lat_DD"), crs=4326)


# Loxgb Ethiopia shapefile
Eth <- sf::st_read("input/eth_adm_2021_shp", layer = "eth_admbnda_adm0_csa_bofedb_itos_2021", quiet = TRUE) %>%
  st_transform(4326)
Eth_regions<- sf::st_read("input/eth_adm_2021_shp", layer = "eth_admbnda_adm1_csa_bofedb_2021", quiet = TRUE) %>%
  st_transform(4326)

p<-ggplot() +
  geom_sf(data = Eth, fill = NA, color = "black") +
  geom_sf(data = Eth_regions, fill = NA, color = "darkgrey")+
  #geom_sf_label(data = Eth_regions,aes(label=ADM1_EN))+
   geom_sf(data=maize_areas,aes(color=Region,size=GYld_Kgpha), na.rm=T) +
  coord_sf()+
  theme_bw()
ggsave(p, file="output/Eth_maize_yield_map.png", width = 6, height = 4, dpi = 300)
p
```
```{r}
library(tmap)
library(tmaptools)
# Set tmap to view mode
tmap_mode("view")
tm_shape(Eth)+
  tm_borders(col = "black") +
tm_shape(Eth_regions) +
  tm_borders(col = "gray")
tm_shape(maize_areas)+
  tm_dots(col = "GYld_Kgpha") +
  tm_mouse_coordinates()
```

* Yield boxplots
```{r}
df %>% drop_na(GYld_Kgpha) %>% 
  ggplot(aes(factor(Treat),GYld_Kgpha, fill=factor(Treat)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 2, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(size = 10), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Grain Yield (kg/ha)",x="Treatments")
```

* Explore yield differences with recNP
```{r}
all<-df %>% group_by(District,Treat) %>%
summarise(mean_diff_all=mean(diff_all,na.rm = T)) %>% drop_na() #%>%
#mutate(mean_diff_all=mean_diff_all * (-1))

```
```{r}
all %>%
ggplot(aes(x=mean_diff_all,y=District,fill=factor(Treat))) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. all yield (%)",y="District")

```

## Analysis of Variance (ANOVA)

Let's check the assumption
```{r}
# Homogeinty of variance
bartlett.test(GYld_Kgpha ~ Treat, data = df)
```
p<0.05 so we can accept the alternative hypothesis which stipulate there are significant differences of variances accross treatments.

* Robust ANOVA
```{r}
# Perform robust ANOVA using Welch's test
welch_result <- oneway.test(GYld_Kgpha ~ Treat, data = df, var.equal = FALSE)
print(welch_result)
```
p of the treatment is statistically significant (p < 0.05), it is likely that treatment type does have a significant effect on average crop yield. In fact there is statistically-significant difference in average crop yield according to treatment.
