---
title: "Maize"
author: "Adama NDOUR"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(gtsummary)
library(inspectdf)
library(sf)
```

## Import data
```{r}
df<- readxl::read_excel("input/Teff_ADAMA.xlsx",sheet = "Sheet1")
```

## Overview of the data
```{r}
str(df)
```
## Data manipulation
```{r warning=FALSE}
# Rename some variables
df<- df %>% rename_with(
   ~ c("Gyld_kgha","recNP_diff","all_diff","soil_type"),.cols= c("Gyld","(RN-trt)/RNP*100", "MYD (All-trt)/All)*100",
  "Soil type-COW"))

# Convert to numeric
df <- df %>% mutate(across(c("Gyld_kgha","recNP_diff","all_diff"),as.numeric))

# Convert string to factor
df<- df %>% mutate_if(is.character,as.factor)

#Remove NA values for yield and soil type
df<- df %>% 
  #drop_na(soil_type) %>% 
  drop_na(Gyld_kgha)

```

## Inspect data
* Categorical variables
```{r}
df %>% select_if(is.character) %>% skim()
```
```{r}
df %>% inspect_cat()
```

```{r}
# Replace NA values in STP
#df %>% mutate(STP=str_replace_na("vertisols")) %>% #filter(ID=="DZ-EAIR_2") %>% view()
df<-df %>% mutate(STP=str_replace_na("vertisols"))
```


```{r}
# numeric variables
df %>% select_if(is.numeric) %>% skim() %>% 
  knitr::kable()
```


## Descriptive statistics
* Yield per treatment
```{r message=FALSE, warning=FALSE}
df %>% select(c("Treat","Gyld_kgha")) %>% 
  tbl_summary(
    by = Treat,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing_text = "Missing"
  ) %>% modify_spanning_header(all_stat_cols() ~ "**Treatments**") %>% add_p()
```

## Explory data analysis
* Map
```{r}
maize_areas<- st_as_sf(df %>% select(c("Lat_DD","Long_DD",
                                      "Treat","GYld_Kgpha","Region")) %>% drop_na() %>% filter(GYld_Kgpha>0,Treat==9),
                        coords=c("Long_DD","Lat_DD"), crs=4326)


# Loxgb Ethiopia shapefile
Eth <- sf::st_read("input/eth_adm_2021_shp", layer = "eth_admbnda_adm0_csa_bofedb_itos_2021", quiet = TRUE) %>%
  st_transform(4326)
Eth_regions<- sf::st_read("input/eth_adm_2021_shp", layer = "eth_admbnda_adm1_csa_bofedb_2021", quiet = TRUE) %>%
  st_transform(4326)

p<-ggplot() +
  geom_sf(data = Eth, fill = NA, color = "black") +
  geom_sf(data = Eth_regions, fill = NA, color = "darkgrey")+
  #geom_sf_label(data = Eth_regions,aes(label=ADM1_EN))+
   geom_sf(data=maize_areas,aes(color=Region,size=GYld_Kgpha), na.rm=T) +
  coord_sf()+
  theme_bw()
ggsave(p, file="output/Eth_maize_yield_map.png", width = 6, height = 4, dpi = 300)
p
```
```{r}
library(tmap)
library(tmaptools)
# Set tmap to view mode
tmap_mode("view")
tm_shape(Eth)+
  tm_borders(col = "black") +
tm_shape(Eth_regions) +
  tm_borders(col = "gray")
tm_shape(maize_areas)+
  tm_dots(col = "GYld_Kgpha") +
  tm_mouse_coordinates()
```

* Yield boxplots
```{r}
df %>% drop_na(Gyld_kgha) %>% 
  ggplot(aes(factor(Treat),Gyld_kgha, fill=factor(Treat)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 2, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(size = 10), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Grain Yield (kg/ha)",x="Treatments")
```
## Analysis of Variance (ANOVA)

Let's check the assumption
```{r}
# Homogeinty of variance
bartlett.test(Gyld_kgha ~ Treat, data = df)
```
p<0.05 so we can accept the alternative hypothesis which stipulate there are significant differences of variances accross treatments.

* Is there statistical significant differences of yield medians accross regions and Regions?
```{r}

# Perform robust ANOVA using Welch's test
oneway.test(Gyld_kgha ~ Region, data = df, var.equal = FALSE)
```
```{r}
# mean differences accross regions and soil type?
df["Region.Soil"]=interaction(df$Region,df$soil_type)
oneway.test(Gyld_kgha ~ Region.Soil, data = df, var.equal = FALSE)
```
## Exploration yield penality 
```{r}
# Recode Treatment
df$Treat<- factor(df$Treat)
df<-df %>% mutate(Treat=case_when(
  Treat=="1" ~ "All",
  Treat=="2" ~ "All-B",
  Treat=="3" ~ "All-Zn",
  Treat=="4" ~ "All-S",
  Treat=="5" ~ "All-K",
  Treat=="6" ~ "All-P",
  Treat=="7" ~ "All-N",
  Treat=="8" ~ "All-KSZnB",
  Treat=="9" ~ "Control",
  Treat=="10" ~ "recNP+S2",
))

# Replace all values for treat1 by 0 in all_diff
#ids<- df %>% filter(Treat==1) %>% pull(ID) %>% unique()
df<-df %>% mutate(all_diff=case_when(
  Treat==1 ~ 0,
  TRUE ~ all_diff
))

# Replace all values for treat8 by 0 in all_recNP
#ids<- df %>% filter(Treat==1) %>% pull(ID) %>% unique()
df<-df %>% mutate(all_diff=case_when(
  Treat==8 ~ 0,
  TRUE ~ all_diff
))
```
### Yield differences with all
* Per region
```{r message=TRUE, warning=FALSE}
all_reg<-df %>% group_by(Region,Treat) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na() %>%
mutate(mean_diff_all=mean_diff_all * (-1))

```
```{r}
p_reg<- all_reg %>%
ggplot(aes(x=mean_diff_all,y=Region,fill=Treat)) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. all yield (%)",y="Region")
p_reg
ggsave(p_reg, file="output/teff/yield_diff_all_region.png", width = 6, height = 4, dpi = 300)
```


* Per region and soil type
```{r}
#df<- df %>% drop_na(soil_type)
all_reg_stp<-df %>% group_by(Region,soil_type,Treat) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na() %>% mutate(mean_diff_all=mean_diff_all * (-1))
all_reg_stp
```
```{r}
p_reg_stp<-all_reg_stp %>%
ggplot(aes(x=mean_diff_all,y=Region,fill=Treat)) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. all yield (%)",y="Region") +
  facet_wrap(~soil_type)
p_reg_stp
ggsave(p_reg_stp, file="output/teff/yield_diff_all_region_stp.png", width = 8, height = 6, dpi = 300)
```




* Per region and moisture domain
```{r}
#df<- df %>% drop_na(MD)
all_reg_md<-df %>% group_by(Region,MD,Treat) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na() %>% mutate(mean_diff_all=mean_diff_all * (-1))
all_reg_md
```
```{r}
p_reg_md<-all_reg_md %>%
ggplot(aes(x=mean_diff_all,y=Region,fill=Treat)) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. all yield (%)",y="Region") +
  facet_wrap(~MD)
p_reg_md
ggsave(p_reg_md, file="output/teff/yield_diff_all_region_md.png", width = 8, height = 6, dpi = 300)
```



### Yield differences with recNP
* Per region
```{r message=TRUE, warning=FALSE}
recNP_reg<-df %>% group_by(Region,Treat) %>%
summarise(mean_diff_recNP=mean(recNP_diff,na.rm = T)) %>% drop_na() %>%
mutate(mean_diff_recNP=mean_diff_recNP * (-1))
recNP_reg
```
```{r}
p_reg<- recNP_reg %>%
ggplot(aes(x=mean_diff_recNP,y=Region,fill=Treat)) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. recNP yield (%)",y="Region")
p_reg
ggsave(p_reg, file="output/teff/yield_diff_recNP_region.png", width = 6, height = 4, dpi = 300)
```


* Per region and soil type
```{r}
df<- df %>% drop_na(soil_type)
recNP_reg_stp<-df %>% group_by(Region,soil_type,Treat) %>%
summarise(mean_diff_recNP=mean(recNP_diff,na.rm = T)) %>% drop_na() %>% mutate(mean_diff_recNP=mean_diff_recNP * (-1))
recNP_reg_stp
```
```{r}
p_reg_stp<-recNP_reg_stp %>%
ggplot(aes(x=mean_diff_recNP,y=Region,fill=Treat)) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. recNP yield (%)",y="Region") +
  facet_wrap(~soil_type)
p_reg_stp
ggsave(p_reg_stp, file="output/teff/yield_diff_recNP_region_stp.png", width = 8, height = 6, dpi = 300)
```




* Per region and moisture domain
```{r}
#df<- df %>% drop_na(MD)
recNP_reg_md<-df %>% group_by(Region,MD,Treat) %>%
summarise(mean_diff_recNP=mean(recNP_diff,na.rm = T)) %>% drop_na() %>% mutate(mean_diff_recNP=mean_diff_recNP * (-1))
recNP_reg_md
```
```{r}
p_reg_md<-recNP_reg_md %>%
ggplot(aes(x=mean_diff_recNP,y=Region,fill=Treat)) +
geom_col()+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1)+
theme_bw()+
  labs(x="Contribution to av. recNP yield (%)",y="Region") +
  facet_wrap(~MD)
p_reg_md
ggsave(p_reg_md, file="output/teff/yield_diff_recNP_region_md.png", width = 8, height = 6, dpi = 300)
```