---
title: "Maize"
author: "Adama NDOUR"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(gtsummary)
library(inspectdf)
library(sf)
library(tidymodels) # ML using tidyverse principles
library(doParallel) # parallel computing
library(recipeselectors) # feature selection
tidymodels_prefer() # handle conflicts between tidymodels and other package
library(bestNormalize) # best normalizing transformation
library(vip) # MLinterpretation
library(DALEX) # MLinterpretation
library(DALEXtra) # MLinterpretation
library(caret) # ML algos
```

## Import data
```{r}
df<- readxl::read_excel("input/Wheat_covariates.xls")
```

## Overview of the data
```{r}
str(df)
```
## Data manipulation
```{r}
# Rename some variables
df<- df %>% rename_with(
   ~ c("recNP_diff","all_diff"),.cols= c("RNP_Per_Yield_diff", "All_Per_Yield_diff"))

# Convert to numeric
#df <- df %>% mutate(across(c("GYld_Kgpha","all_tr1","recNP_tr8"),
                           #as.numeric))
```

## Inspect data
* Categorical variables
```{r}
df %>% select_if(is.character) %>% skim()
```
```{r}
df %>% select(Region, District, Farmer_name,Soil_table) %>%  inspect_cat()
```


```{r}
# numeric variables
df %>% select_if(is.numeric) %>% skim() %>% 
  knitr::kable()
```


## Descriptive statistics
* Yield per Trtment
```{r message=FALSE, warning=FALSE}
df %>% select(c("Trt","AGY")) %>% 
  tbl_summary(
    by = Trt,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing_text = "(Missing)"
  ) %>% modify_spanning_header(all_stat_cols() ~ "**Trtments**") %>% add_p()
```

## Explory data analysis
* Map
```{r}
wheat_areas<-df %>% select(c("Lat_DD","Lon_DD",
    "Trt","AGY","Farmer_name")) %>% drop_na() %>% filter(AGY>0) %>% group_by(Farmer_name) %>% 
  summarise(Lon=mean(Lon_DD),Lat=mean(Lat_DD),
            AGY=mean(AGY))

wheat_areas<- st_as_sf(wheat_areas,
                        coords=c("Lon","Lat"), crs=4326)


# Loxgb Ethiopia shapefile
Eth <- sf::st_read("input/eth_adm_2021_shp", layer = "eth_admbnda_adm0_csa_bofedb_itos_2021", quiet = TRUE) %>%
  st_transform(4326)
Eth_regions<- sf::st_read("input/eth_adm_2021_shp", layer = "eth_admbnda_adm1_csa_bofedb_2021", quiet = TRUE) %>%
  st_transform(4326)

p<-ggplot() +
  geom_sf(data = Eth, fill = NA, color = "black") +
  geom_sf(data = Eth_regions, fill = NA, color = "darkgrey")+
  #geom_sf_label(data = Eth_regions,aes(label=ADM1_EN))+
   geom_sf(data=wheat_areas,aes(color=AGY), na.rm=T) +
  scale_color_gradient2(low = "blue",high = "red")+
  coord_sf()+
  theme_bw()
ggsave(p, file="output/wheat/Eth_wheat_yield_map.png", width = 6, height = 4, dpi = 300)
p
```
```{r}
library(tmap)
library(tmaptools)
# Set tmap to view mode
tmap_mode("view")
tm_shape(Eth)+
  tm_borders(col = "black") +
tm_shape(Eth_regions) +
  tm_borders(col = "gray")
tm_shape(wheat_areas)+
  tm_dots(col = "AGY") +
  tm_mouse_coordinates()
```

* Yield boxplots
```{r}
# Recode Trtment
df$Trt<- factor(df$Trt)
df<-df %>% mutate(Trt=case_when(
  Trt=="1" ~ "All",
  Trt=="2" ~ "All-B",
  Trt=="3" ~ "All-Zn",
  Trt=="4" ~ "All-S",
  Trt=="5" ~ "All-K",
  Trt=="6" ~ "All-P",
  Trt=="7" ~ "All-N",
  Trt=="8" ~ "All-KSZnB",
  Trt=="9" ~ "Control",
  Trt=="10" ~ "recNP+S2",
))

```


```{r}
p_agy<-df %>% drop_na(AGY) %>% 
  ggplot(aes(factor(Trt),AGY, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 2, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(size = 10), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Grain Yield (kg/ha)",x="Treatments")
p_agy
ggsave(p_agy, file="output/wheat/AGY_national.png", width = 6, height = 4, dpi = 300)
```


* Yield boxplots per soil type
```{r}
p_agy_stp<-df %>% drop_na(AGY) %>% 
  ggplot(aes(factor(Trt),AGY, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size = 8), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Grain Yield (kg/ha)",x="Treatments")+
  facet_wrap(~Soil_table)
p_agy_stp
ggsave(p_agy_stp, file="output/wheat/AGY_national_stp.png", width = 8, height = 6, dpi = 300)
```
* Yield boxplots per moisture domain
```{r}
p_agy_MD<-df %>% drop_na(AGY) %>% 
  ggplot(aes(factor(Trt),AGY, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size = 8), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Grain Yield (kg/ha)",x="Treatments")+
  facet_wrap(~MD)
p_agy_MD
ggsave(p_agy_MD, file="output/wheat/AGY_national_MD.png", width = 8, height = 6, dpi = 300)
```

*  Biomass boxplots
```{r}
df$ABY<- df$ABY * 1000
```

```{r}
p_aby<-df %>% drop_na(ABY) %>% filter(ABY<25000) %>% 
  ggplot(aes(factor(Trt),ABY, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 2, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(size = 10), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Biomass (kg/ha)",x="Treatments")
p_aby
ggsave(p_aby, file="output/wheat/ABY_national.png", width = 6, height = 4, dpi = 300)
```

* Biomass boxplots per soil type
```{r}
p_aby_stp<-df %>% drop_na(ABY) %>% filter(ABY<25000) %>% 
  ggplot(aes(factor(Trt),ABY, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size = 6), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Biomass (kg/ha)",x="Treatments")+
  facet_wrap(~Soil_table)
p_aby_stp
ggsave(p_aby_stp, file="output/wheat/ABY_national_stp.png", width = 8, height = 6, dpi = 300)
```

* Biomass boxplots per moisture domain
```{r}
p_aby_MD<-df %>% drop_na(ABY) %>% filter(ABY<25000) %>% 
  ggplot(aes(factor(Trt),ABY, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size = 8), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="Biomass (kg/ha)",x="Treatments")+
  facet_wrap(~MD)
p_aby_MD
ggsave(p_aby_MD, file="output/wheat/ABY_national_MD.png", width = 8, height = 6, dpi = 300)
```


*  HI boxplots

```{r}
p_hi<-df %>% drop_na(HI)  %>% 
  ggplot(aes(factor(Trt),HI, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 2, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(size = 10), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="HI",x="Treatments")
p_hi
ggsave(p_hi, file="output/wheat/HI_national.png", width = 6, height = 4, dpi = 300)
```

* HI boxplots per soil type
```{r}
p_hi_stp<-df %>% drop_na(HI)  %>% 
  ggplot(aes(factor(Trt),HI, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size = 6), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="HI",x="Treatments")+
  facet_wrap(~Soil_table)
p_hi_stp
ggsave(p_hi_stp, file="output/wheat/HI_national_stp.png", width = 8, height = 6, dpi = 300)
```

* HI boxplots per moisture domain
```{r}
p_hi_MD<-df %>% drop_na(HI)  %>% 
  ggplot(aes(factor(Trt),HI, fill=factor(Trt)))+
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 21, size = 1, color = "white",fill="blue")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=90,size = 8), 
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.position = 'none') +
  labs(y="HI",x="Treatments")+
  facet_wrap(~MD)
p_hi_MD
ggsave(p_hi_MD, file="output/wheat/HI_national_MD.png", width = 8, height = 6, dpi = 300)
```

## Analysis of Variance (ANOVA)

Let's check the assumption
```{r}
# Homogeinty of variance
bartlett.test(AGY ~ Trt, data = df)
```

```{r}

# Perform robust ANOVA using Welch's test
oneway.test(all_diff ~ Region, data = df, var.equal = FALSE)
```
```{r}
# mean differences accross regions and soil type?
df["Region.Soil"]=interaction(df$Region,df$Soil_table)
oneway.test(AGY ~ Region.Soil, data = df, var.equal = FALSE)
```

## Exploration yield penality 

### Yield differences with all

* National level
```{r}
all_nat<- df %>% group_by(Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na()
all_nat
write.csv(all_nat,"output/wheat/all_diff_nat.csv", row.names = F)
```
```{r}
colors<-c("#bdc9e1","#cb181d","#bae4b3","#969696","#df65b0",
          "#6a51a3","#fecc5c","#d7b5d8","#525252","#238b45")
p_nat_all<- all_nat %>%
ggplot(aes(x=mean_diff_all,y=Trt,fill=Trt)) +
geom_col(width = 0.5)+
  scale_fill_manual(values = colors)+
  #scale_y_discrete(breaks = seq(0, 25, by = 5))+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x="Average Yield Penalty [all] (%)",fill="Treatment",
       y="")
p_nat_all
ggsave(p_nat_all, file="output/wheat/yield_diff_all_national.png", width = 6, height = 4, dpi = 300)
```

* National level per soil
```{r}
all_nat_soil_table<- df %>% group_by(Soil_table,Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na()
all_nat_soil_table
write.csv(all_nat_soil_table,"output/wheat/all_diff_nat_soil_table.csv", row.names = F)
```
```{r}
colors<-c("#bdc9e1","#cb181d","#bae4b3","#969696","#df65b0",
          "#6a51a3","#fecc5c","#d7b5d8","#525252","#238b45")
p_all_nat_soil_table<- all_nat_soil_table %>%
ggplot(aes(x=mean_diff_all,y=Trt,fill=Trt)) +
geom_col(width = 0.5)+
  scale_fill_manual(values = colors)+
  #scale_y_discrete(breaks = seq(0, 25, by = 5))+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x=" Average yield penalty [all] (%)",fill="Treatment",y="")+
  facet_wrap(~Soil_table) +
  theme(axis.text.y = element_text(size = 4))
p_all_nat_soil_table
ggsave(p_all_nat_soil_table, file="output/wheat/yield_diff_all_national_soil_table.png", width = 6, height = 4, dpi = 300)
```
* National level per moisture domain
```{r}
all_nat_MD<- df %>% group_by(MD,Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na()
all_nat_MD
write.csv(all_nat_MD,"output/wheat/all_diff_nat_MD.csv", row.names = F)
```
```{r}
colors<-c("#bdc9e1","#cb181d","#bae4b3","#969696","#df65b0",
          "#6a51a3","#fecc5c","#d7b5d8","#525252","#238b45")
p_all_nat_MD<- all_nat_MD %>%
ggplot(aes(x=mean_diff_all,y=Trt,fill=Trt)) +
geom_col(width = 0.5)+
  scale_fill_manual(values = colors)+
  #scale_y_discrete(breaks = seq(0, 25, by = 5))+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x=" Average yield penalty [all] (%)",fill="Treatment",y="")+
  facet_wrap(~MD) +
  theme(axis.text.y = element_text(size = 8))
p_all_nat_MD
ggsave(p_all_nat_MD, file="output/wheat/yield_diff_all_national_MD.png", width = 6, height = 4, dpi = 300)
```




* Per region
```{r message=TRUE, warning=FALSE}
all_reg<-df %>% group_by(Region,Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) %>% drop_na()
write.csv(all_reg,"output/wheat/all_diff_reg.csv", row.names = F)
```
```{r}
colors<-c("#bdc9e1","#cb181d","#bae4b3","#969696","#df65b0",
          "#6a51a3","#fecc5c","#d7b5d8","#525252","#238b45")
p_reg<- all_reg %>%
ggplot(aes(x=mean_diff_all,y=Region,fill=Trt)) +
geom_col(width = 0.5)+
  scale_fill_manual(values = colors)+
  #scale_y_discrete(breaks = seq(0, 25, by = 5))+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x="Average Yield Penalty [all] (%)",y="Region")
p_reg
ggsave(p_reg, file="output/wheat/yield_diff_all_region.png", width = 6, height = 4, dpi = 300)
```


* Per region and soil type
```{r}
#df<- df %>% drop_na(Soil_table)
all_reg_soil_table<-df %>% group_by(Region,Soil_table,Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) 
all_reg_soil_table
write.csv(all_reg_soil_table,"output/wheat/all_diff_reg_soil_table.csv", row.names = F)
```
```{r}
p_reg_soil_table<-all_reg_soil_table %>%
ggplot(aes(x=mean_diff_all,y=Region,fill=Trt)) +
geom_col()+
scale_fill_manual(values = colors)+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x="Average Yield Penalty [all] (%)",y="Region") +
  facet_wrap(~Soil_table)
p_reg_soil_table
ggsave(p_reg_soil_table, file="output/wheat/yield_diff_all_region_soil_table.png", width = 8, height = 6, dpi = 300)
```

* Per region and moisture domain
```{r}
#df<- df %>% drop_na(MD)
all_reg_md<-df %>% group_by(Region,MD,Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) 
all_reg_md
write.csv(all_reg_md,"output/wheat/all_diff_reg_md.csv", row.names = F)
```
```{r}
p_reg_md<-all_reg_md %>%
ggplot(aes(x=mean_diff_all,y=Region,fill=Trt)) +
geom_col(width = 0.5)+
scale_fill_manual(values = colors)+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x="Average Yield Penalty [all] (%)",y="Region") +
  facet_wrap(~MD)
p_reg_md
ggsave(p_reg_md, file="output/wheat/all_diff_region_md.png", width = 8, height = 6, dpi = 300)
```



* Per region and district
```{r}
#df<- df %>% drop_na(Soil_table)
all_reg_district<-df %>% group_by(Region,District,Trt) %>%
summarise(mean_diff_all=mean(all_diff,na.rm = T)) 
all_reg_district
write.csv(all_reg_district,"output/wheat/all_diff_reg_district.csv", row.names = F)
```
```{r}
p_reg_district<-all_reg_district %>% filter(District !="Sankura") %>% 
ggplot(aes(x=mean_diff_all,y=District,fill=Trt)) +
geom_col()+
scale_fill_manual(values = colors)+
geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 0.6)+
theme_bw()+
  labs(x="Average Yield Penalty [all] (%)",y="Region",
       fill="Treatment") +
  facet_wrap(~Region)+
  theme(axis.text.y = element_text(size=4),
        legend.text = element_text(size=6))
p_reg_district
ggsave(p_reg_district, file="output/wheat/yield_diff_all_region_district.png", width = 8, height = 6, dpi = 300)
```
## ML modeling
* Load data
```{r}
# Read data
dfcov<- readxl::read_excel("input/wheat_ydiff_covariates.xls", sheet = "Cleaned")
```

* Select yield and covariates
```{r}
dfcov<- dfcov %>% select("All_Yield_diff":"Mo_rf")
```

* Check the yield distribution
```{r}
p<-ggplot(dfcov, aes(All_Yield_diff)) +
  geom_histogram(fill="#ffa500",color="white") +
  theme_bw()+
  labs(x="Grain Yield Difference [all](kg/ha)")
ggsave(p, file="output/wheat/Eth_wheat_hist.png", width = 4, height = 3, dpi = 300)
p
```

* Preprocessing
```{r}
# Split data to create partitions
#3df<- df %>% select(-c(nfert_kgha,seed_rate_kgha))
df_split <- initial_split(dfcov, prop = 0.8)
df_train <- training(df_split)
df_test <- testing(df_split)
```

Let's build a recipe
```{r message=FALSE, warning=FALSE}
recipe <- recipe(All_Yield_diff ~ ., data = df_train) %>% # filtering
  #step_naomit() %>% 
  # Apply Yeo-Johnson transformation for left-skewed yield data
  #step_YeoJohnson(cc_yield_kgha) %>% 
  # Convert string to factor
  #step_string2factor(all_nominal()) %>%
  # remove no variance predictors which provide no predictive information 
  #recipes::step_nzv(all_nominal()) %>%
  # Encode factor as dummy variables
  #step_dummy(all_nominal(),one_hot = T) %>%
  #step_lencode_mixed(all_nominal(), outcome=vars(cc_yield_kgha)) %>% 
  # Replace na values by mean
  step_impute_mean(all_predictors(),all_outcomes()) %>% 
  # Remove non-variance variables
  step_nzv(where(is.numeric)) %>%
  # Normalize numeric variables
  step_orderNorm(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  # remove any linear combinations
  #step_pca(all_predictors(), threshold = .9) %>% 
  prep()
```

```{r}
recipe %>% juice() %>% dim()
```

* Recursive feature elimination
```{r}
registerDoParallel() # Enable distributed computation
train_set <- juice(recipe)
test_set <- bake(recipe, new_data = df_test)
# define five-fold cross-validation scheme
control_feature <- rfeControl(functions=rfFuncs, method="cv", number=5)
#
# recursive feature elimination; 'sizes' stands for the number of variables
x=train_set|>select(-All_Yield_diff)
y=train_set$All_Yield_diff
feature_selection <- rfe(x=x,y=y,sizes = c(1:28), rfeControl=control_feature)
#
# rmse vs. number of variables (ordered from very to not important)
trellis.par.set(caretTheme())
plot(feature_selection, type = c("g", "o"))
#ggsave(file="output/rmse_rfe.png",plot = last_plot(), width = 5, height = 4, dpi = 300)


```
```{r}
r2_vals<- as_tibble(feature_selection$results$Rsquared) %>% mutate(Variables=row_number())
mx<- max(r2_vals$value)
idx_mx<- which(r2_vals==mx)
p<-ggplot(data=r2_vals, aes(Variables,y=value)) + 
  geom_line()+
  geom_point(fill="white", color="blue",size=2, alpha=0.5) +
  geom_point(aes(x=idx_mx,y=mx),color="red",size=2)+
  theme_bw() +
  labs(y="R2 (Cross-Validation)")
ggsave(p,file="output/r2_rfe.png", width = 5, height = 4, dpi = 300)
p

```

```{r}
# list all variables
list_of_var <- feature_selection$optVariables
list_of_var <- append("cc_yield_kgha", list_of_var)
#
# select 10 most important variables only
train_data <- train_set[,list_of_var]
test_data <- test_set[,list_of_var]
```

* Define models
```{r}
# Random forest
rf_spec <- rand_forest(
  trees = tune(),
  mtry = tune(),
  min_n = tune()
) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("regression")
rf_grid<- grid_regular(
  mtry(c(4,12)), 
  trees(),
 min_n(),
 levels = 8
)


# eXtreme Grxgbient Boost
xgb_spec <- boost_tree(
  trees = tune(),
  mtry = tune(),
  learn_rate = tune()
) %>%
  set_engine("xgboost", importance="permutation") %>%
  set_mode("regression")
xgb_grid<- grid_regular(
  mtry(c(4,12)), 
  trees(),
 learn_rate(),
 levels = 8
)
```

* Define resampling and metrics
```{r}
cv <- vfold_cv(juice(recipe), v = 5)
set.seed(6735)
#cv <- vfold_cv(juice(prep(recipe_wf)), v = 5)
#cv <- vfold_cv(train_data, v = 5)
metrics <- metric_set(rmse,rsq)
```

* Train and evaluate the models

```{r}
registerDoParallel() # Enable distributed computation
# RF
rf_res <- tune_grid(
  rf_spec,
  All_Yield_diff ~ .,
  cv,
  metrics = metrics,
  grid=rf_grid)

rf_res %>%
  collect_metrics()
```

```{r}
rf_res %>%
  tune::show_best(metric = "rsq") %>%
  knitr::kable()
```
```{r}
rf_res %>% autoplot()
```
